Collocated Deduplication with Fault Isolation for Virtual Machine Snapshot Backup

### 2013-06-25 snapshot_system/abs.tex

\begin{abstract}
A cloud environment that  hosts a large number of  virtual machines (VMs) has
a high storage demand for frequent backup of system image snapshots.
Deduplication of data blocks can lead a big reduction of redundant blocks when
their signatures are identical. However it is expensive and less fault-resilient to perform a global deduplication
using signatures and let a data block share by many virtual machines. 
This paper studies a cluster-based VM-centric scheme which collocates a lightweight
backup service with the cloud service and it integrates multiple duplicate detection strategies that  
localize the deduplication as much as possible within each virtual machine.
It also organizes the write of small data chunks into large file system blocks so
that each underlying file block is associated with one VM for most of cases.
Our analysis shows that  this VM centric scheme 
can  provide  better fault tolerance while using a small amount of computing and storage resource. 
This  paper  provides a comparative evaluation of this scheme in  accomplishing a high deduplication 
efficiency while sustaining a good backup throughput.

>>>>

\begin{abstract}
A cloud environment that  hosts a large number of  virtual machines (VMs) has
a high storage demand for frequent backup of system image snapshots.
Deduplication of data blocks can lead a big reduction of redundant blocks when
their signatures are identical. However it is expensive and less fault-resilient to perform a global deduplication
using signatures and let a data block share by many virtual machines. 
This paper studies a VM-centric scheme which collocates a lightweight
backup service with other cloud services in a cluster and it integrates multiple duplicate detection strategies 
that  localize deduplication as much as possible within each virtual machine.
It also organizes the write of small data chunks into large file system blocks so
that each underlying file block is associated with one VM for most of cases.
Our analysis shows that  this VM centric scheme 
can  provide  better fault tolerance while using a small amount of computing and storage resource. 
This  paper  provides a comparative evaluation of this scheme in  accomplishing a high deduplication 
efficiency while sustaining a good backup throughput.

### 2013-06-25 snapshot_system/challenge.tex

As discussed earlier, collocating the backup service on the existing
cloud cluster avoids the extra cost to acquire a dedicated backup facility
and reduces the network bandwidth consumption in transferring the un-deduplicated
raw data for backup. 
%The running of such a service is subject to the resource availability of the system and 
We discuss the design considerations as follows.

@@

\begin{figure}[htb]
    \centering
    \includegraphics[width=3in]{images/colocated-arch.pdf}
    \label{fig-colocated-arch}
    \caption{Collocated VM Backup System}
\end{figure}

\begin{itemize}
\item {\em Collocating with cluster-based cloud services.}
We collocate the backup service with other cloud services in a cluster
and use a distributed file system~\cite{GFS2004,Hadoop} to 
store the backup with multiple replicas. 
There might be a concern that storing many versions of snapshots would
go against the purpose of deduplication.
The previous research shows the deduplication can compress the backup copies 
effectively in a 10:1 or even 15:1 rage. Assume on the average the compression
ratio is $c$:1.  When the number of versions
saved in the backup storage is $l$  on the average, 
the amount of storage space in the cluster allocated  for the backup would be $\frac{l}{l+c}$.
For example, when $l=5$ and $c=10$, 33\% of the cluster storage is used for the backup.
When $l=9$ and $c=13$, storage usage is 40\%. 
That is acceptable given today's cheap cost for disk storage.
In the case of Alibaba, each machine hosts 25 VMs on the average and each VM uses 40GB data.
Each machine however can typically provide over 10TB storage.
Thus the normal disk usage  is about 1TB per machine, and adding 30\% to 100\% 
extra space for the backup is not a practical concern, especially with the
added benefits of cheaper running costs and lower network bandwidth requirements.

Since the file block size in the Hadoop and GFS is uniform and large with 64MB as a default setting,
the content chunk in a deduplication system is of nonuniform size with 4KB or 8KB on average.
We need to build an intermediate layer that supports large snapshot writing with append operations
and infrequent snapshot access.

\item {\em Fault tolerance and isolation.}
The distributed file system~\cite{GoogleFS,Hadoop} provides the replication support
and we can leverage existing open-source software to implement such a layer.
When the number of machines failed exceeds the replication degree, there could be
 some data loss.  Storage deduplication worsens the impact of such faults
because  the majority of each snapshot (e.g. over 90\%) is deduplicated
and the corresponding data chunks stored in the underlying file blocks
are shared by many VMs. 
%The failure of one data chunk or block can impact many sharing VMs. 
In large scale cloud, machine failures happen at daily basis,
the loss of a shared chunks can affect many VM users whose snapshots share this 
data block.

Previous work in deduplication~\cite{extreme_binning09,sparseindex09} has
studied the solutions to reduce memory requirements.
One important issue not considered in the previous work is how content sharing affects
fault isolation.
Because a content chunk is compared with a content signature collected from other users,
this artificially creates data dependency among different VM users.
% since storage for shared identical content chunks can become a failure point.
Without any control of such data sharing, we can only increase  
replication for global datasets to enhance the availability,
but this incurs significantly higher costs.
Loss of a small number of shared data chunks can 
cause the unavailability of snapshots for a large number of virtual machines.
Thus a key goal  in our design
is restricting the impact of  data loss to a limited number of VMs. 
It is not desirable that small scale of data failure affects the backup of many VMs.
%when users access snapshots in a recovery process.

\end{itemize}

%\item {\em Requirement for backup throughput and low computing resource usage}.
Because of collocation of this snapshot service with other existing cloud services, 
%As a large-scale cluster hosting tens of thousand of active VMs everyday,  the amount of data
%to be processed can be huge. 
the computing resources allocated for the snapshot service is limited, especially
because of its lower priority.
The key usage of resource for backup is memory for storing and comparing the fingerprints. 
We will consider the approximation techniques with less memory consumption
studied in~\cite{extreme_binning09,Guo2011} along with the fault isolation consideration discussed below. 
%A snapshot backup service shall not compete cpu, memory, or I/O bandwidth with VMs. 
%Fingerprint-based search does require a significant amount of memory for fingerprint lookup
%and comparison.  This competes memory resource  with the existing VMs.
%Our design is to identify a low-cost and low-profile scheme for deduplication.
%Specifically, the memory usage of snapshot service can never exceed 500MB in any node.
Excessive use of bandwidth and computing resource for backup can create
noticeable contention for resources with the existing cloud, which is not
preferred and may not not be acceptable for production system operation. 
Thus it is desirable that backup for all nodes
can be conducted in parallel and in a short period of time,
and any centralized or cross-machine communication for
deduplication should not become a bottleneck.
For example, in an Aliyun cluster with over 1,000 nodes and each hosts over 25 VMs, 
the system must finish saving daily snapshots of all VMs in a few hours. 
Often a cloud has a few hours of light workload each day (e.g. midnight),  
which creates an small window for automatic backup.
The backup service can generate hundreds of terabytes of data in such a time window
without optimization and thus the resource content is considered carefully as we design
a fault-resilient cluster-based solution. 
%In our typical 1000 nodes cluster, each node hosts 25 VMs, each VM has 40GB of data on average, that translates to backup throughput of 139GB/second, or 
%500TB/hour.  The system must finish saving daily snapshots of all VMs in few hours.

>>>>

As discussed earlier, collocating the backup service on the existing
cloud cluster avoids the extra cost to acquire a dedicated backup facility
and reduces the network bandwidth consumption in transferring the un-deduplicated
raw data for backup. 
Figure~\ref{fig:collocated} illustrates the cluster architeture where
each physical runs a backup service and a distribued file system (DFS)~\cite{GFS2004,Hadoop} 
serves a backup store  for the snapshopts.
The previous study shows that 
 deduplication can compress the backup copies 
effectively in a 10:1 or even 15:1 rage. 
Therefore  the portion of space in a cluster
allocated for snapshots of data should not dominate
the cluster storage usage.

%The running of such a service is subject to the resource availability of the system and

@@

\begin{figure}[htb]
    \centering
    \includegraphics[width=3in]{images/colocated-arch.pdf}
    \label{fig:collocated}
    \caption{Collocated VM Backup System.}
\end{figure}

%{\em Collocating with cluster-based cloud services.}
%We collocate the backup service with other cloud services in a cluster
%and use a distributed file system
%store the backup with multiple replicas. 
%There might be a concern that storing many versions of snapshots would
%go against the purpose of deduplication.

%Assume on the average the compression
%ratio is $c$:1.  When the number of versions
%saved in the backup storage is $l$  on the average, 
%the amount of storage space in the cluster allocated  for the backup would be $\frac{l}{l+c}$.
%For example, when $l=5$ and $c=10$, 33\% of the cluster storage is used for the backup.
%When $l=9$ and $c=13$, storage usage is 40\%. 
%That is acceptable given today's cheap cost for disk storage.
%In the case of Alibaba, each machine hosts 25 VMs on the average and each VM uses 40GB data.
%Each machine however can typically provide over 10TB storage.
%Thus the normal disk usage  is about 1TB per machine, and adding 30\% to 100\% 
%extra space for the backup is not a practical concern, especially with the
%added benefits of cheaper running costs and lower network bandwidth requirements.

We discuss the design considerations as follows. 
\begin{itemize}
\item {\em Deduplication localization, sharing minmization, and fault tolerance.}

%Previous work in deduplication~\cite{extreme_binning09,sparseindex09} has
%studied the solutions to reduce memory requirements.
%One important issue not considered in the previous work is how 
Because a data chunk is compared with signatures collected from other VMs during
the deduplication process, only one copy of duplicates is stored in the backup storage
and this artificially creates data dependency among different VM users. 
Content sharing via deduplication affects fault isolation since machine failures happen at daily 
basis in a large-scale cloud and
%failure of a shared block affects many VMs.
%the loss of a shared chunks can affect many VM users whose snapshots share this 
%data block. 
loss of a small number of shared data chunks can 
cause the unavailability of snapshots for a large number of virtual machines.
Localizing the impact of deduplication can incrase fault isolation and resilience.
Thus from the fault tolerance point of view,  duplicate sharing among mutiple VMs is 
discourage. On the other hand, we need to seek for a tradeoff since there are
a significant number of duplicates cross VMs.

% since storage for shared identical content chunks can become a failure point.
%Without any control of such data sharing, we can only increase  
%replication for global datasets to enhance the availability,
%but this incurs significantly higher costs.
%Thus a key goal  in our design
%is restricting the impact of  data loss to a limited number of VMs. 
%It is not desirable that small scale of data failure affects the backup of many VMs.
%when users access snapshots in a recovery process. 
\item{\bf Packaging  data chunks as file system blocks.}
Since the file block size in the Hadoop and GFS is uniform and large with 64MB as a default setting,
the content chunk in a deduplication system is of nonuniform size with 4KB or 8KB on average.
We need to build an intermediate layer that supports large snapshot writing with append operations
and infrequent snapshot access.
Packaging that maps data chunks to file system blocks can create data dependence among VMs
since a file block can be shared even more VMs.
Thus we need to consider a minimum assoocation of a fie system block to VMs in the packaging process.
%The distributed file system~\cite{GoogleFS,Hadoop} provides the replication support
%and we can leverage existing open-source software to implement such a layer.
%When the number of machines failed exceeds the replication degree, there could be
% some data loss.  
%Storage deduplication worsens the impact of such faults
%because  the majority of each snapshot (e.g. over 90\%) is deduplicated
%and the corresponding data chunks stored in the underlying file blocks
%are shared by many VMs. 
%The failure of one data chunk or block can impact many sharing VMs. 
\end{itemize}

%\item {\em Requirement for backup throughput and low computing resource usage}.
Because of collocation of this snapshot service with other existing cloud services, 
%As a large-scale cluster hosting tens of thousand of active VMs everyday,  the amount of data
%to be processed can be huge. 
cloud providers wish that the backup service only consumes  small resources
with a minimal impact to the existing cloud services.
%the computing resources allocated for the snapshot service is limited, especially
%because of its lower priority.
The key usage of resource for backup is memory for storing and comparing the fingerprints. 
We will consider the approximation techniques with less memory consumption
studied in~\cite{extreme_binning09,Guo2011} along with the fault isolation consideration discussed below. 
%A snapshot backup service shall not compete cpu, memory, or I/O bandwidth with VMs. 
%Fingerprint-based search does require a significant amount of memory for fingerprint lookup
%and comparison.  This competes memory resource  with the existing VMs.
%Our design is to identify a low-cost and low-profile scheme for deduplication.
%Specifically, the memory usage of snapshot service can never exceed 500MB in any node.
%Excessive use of bandwidth and computing resource for backup can create
%noticeable contention for resources with the existing cloud, which is not
%preferred and may not not be acceptable for production system operation. 
%Thus it is desirable that backup for all nodes
%can be conducted in parallel and in a short period of time,
%and any centralized or cross-machine communication for
%deduplication should not become a bottleneck.
%For example, in an Aliyun cluster with over 1,000 nodes and each hosts over 25 VMs, 
%the system must finish saving daily snapshots of all VMs in a few hours. 
%Often a cloud has a few hours of light workload each day (e.g. midnight),  
%which creates an small window for automatic backup.
%The backup service can generate hundreds of terabytes of data in such a time window
%without optimization and thus the resource content is considered carefully as we design
%a fault-resilient cluster-based solution. 
%In our typical 1000 nodes cluster, each node hosts 25 VMs, each VM has 40GB of data on average, that translates to backup throughput of 139GB/second, or 
%500TB/hour.  The system must finish saving daily snapshots of all VMs in few hours.
